---
title: "NYC Tree Canopy and Forestry Service Requests Analysis"
output: html_document
---
By Jessica Saab  
For NYU Spatial Analysis and Visualization, Fall 2021  

### Loading libraries
```{r, collapse=TRUE}
library(tidyverse)
library(fs)
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
library(skimr)
library(ggthemes)

# bringing in a raw file and creating a clean version
calls_raw <- read_csv("Forestry_Service_Requests.csv", 
                      col_types = cols(CommunityBoard = col_character(), 
                                       Postcode = col_character(), 
                                       InitiatedDate = col_datetime(format = "%m/%d/%Y %H:%M:%S"),
                                       ClosedDate = col_datetime(format = "%m/%d/%Y %H:%M:%S"),
                                       )
                      ) %>%  clean_names

canopy <- read_csv("ny4p_tree_canopy_coverage.csv") %>% clean_names %>% rename("cb_name" = cb)
```

### Creating a main dataset

My first dataset catalogs Forestry Service Requests to the NYC Parks Department.  
Source: https://data.cityofnewyork.us/Environment/Forestry-Service-Requests/mu46-p9is

I first took a look at the data to select some columns for analysis.  
With over 650,000 obervations and 47 columns, there was a lot of data to look at.  
I selected seven columns, renamed them, filtered out irregular values, and created some new columns with values.  
Then I selected a subset and arranged them by date.

My second dataset catalogs tree canopy cover in NYC.  
Source: http://www.ny4p.org/data-and-research/research-library/open-space-profiles  
It was aggregated by New Yorkers for Parks, a nonprofit organization that advocates for and researches open space in NYC. The dateset is not available to the public but I selected a specific measurement from their 2021 Open Space Profiles, which are available as PDFs, and created a simple CSV with the information on tree canopy cover.  
```{r}
calls <- select(calls_raw, # selecting columns from the full dataset
  sr_status, # whether the case is open or closed
  sr_category, # type of forestry request or concern 
  sr_priority, # priority level given by agency
  borough_code, # borough name
  community_board, # community board
  initiated_date, # start date
  closed_date # end date
  ) %>%
  rename( # renaming columns
    "status" = sr_status, 
    "category" = sr_category, 
    "priority" = sr_priority, 
    "cb" = community_board, 
    "start_date" = initiated_date, 
    "end_date" = closed_date
    ) %>% 
  filter(str_length(cb) == 3) %>% # filtering out community board codes that are not three digits
  filter(!is.na(cb)) %>% # filtering out community boards that are NULL
  filter(!is.na(start_date)) %>% # filtering out start dates that are NULL
  filter(!is.na(end_date)) %>% # filtering out end dates that are NULL which are open cases
  filter(category != "Remove Debris") %>% #only two years of data
  filter(category != "Remove Stump") %>% #only two years of data
  mutate( # making new columns
    boro = str_sub(cb, 1, 1), # taking the first digit to find the boro
    district = str_sub(cb, 2, 3), # taking the last two digits to find the community board
    start = (str_sub(start_date, 1, 7)),
    end = (str_sub(end_date, 1, 7)),
    boro_initials = recode(
      boro, "1" = "MN", "2" = "BX", "3" = "BK", "4" = "QN", "5" = "SI"),
    boro_name = recode(
      boro, "1" = "Manhattan", "2" = "Bronx", "3" = "Brooklyn", "4" = "Queens", "5" = "Staten Island"),
    cb_name = str_c(boro_initials, district)
  ) %>% 
  select(boro_name, cb_name, start, end, status, priority, category) %>% 
  arrange(start)
calls <- left_join(calls, canopy, "cb_name" = "cb_name") # joining my two datasets together
calls
```

### Analysis: Tree canopy coverage in NYC
Finding the average of tree canopy coverage by borough
```{r}
canopy_nyc <- calls %>% 
  group_by(boro_name) %>% 
  summarise(canopy_cover = mean(canopy_cover, na.rm = TRUE))
canopy_nyc
```

### Plot: Ranking of tree canopy coverage in NYC boroughs
```{r}
canopy_nyc_plot <- ggplot(canopy_nyc, aes(reorder(boro_name, canopy_cover), canopy_cover, fill = boro_name)) +
  geom_col(size = 1, alpha = 0.9) +
  coord_flip() +
  scale_y_continuous() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(), axis.title.y = element_blank()) +
  theme(legend.position = "NONE") +
  labs(
    title = "Tree Canopy Coverage by NYC Borough",
    y = "Total proportion of tree canopy coverage",
    caption = "Source: New Yorkers for Parks"
  ) +
  scale_fill_brewer(palette = "Set2")
canopy_nyc_plot
```

### Analysis: Finding out total service calls by borough
Finding total number of calls by borough, excluding open cases, by grouping by year and borough, and sorting by number of calls.
```{r}
calls_boro <- calls %>% 
  mutate(year = str_sub(start, 1, 4)) %>% 
  group_by(year, boro_name) %>% 
  summarise(total_calls = n()) %>% 
  arrange(desc(total_calls)) %>% 
  ungroup()
calls_boro
```

### Analysis: Finding out total types of service calls by borough
Finding total number of types of calls by borough, excluding open cases, by grouping by year and borough and category, and sorting by number of calls.
```{r}
call_types_boro <- calls %>% 
  mutate(year = str_sub(start, 1, 4)) %>% 
  group_by(year, boro_name, category) %>% 
  summarise(total_calls = n()) %>% 
  arrange(desc(total_calls)) %>% 
  ungroup()
call_types_boro
```

### Plot: Calls by Borough in all NYC
Creating a bar graph of total calls by borough.
```{r}
calls_boro_plot <- ggplot(calls_boro, aes(year, total_calls, group = boro_name, fill = boro_name)) +
  geom_col(size= 1, alpha = 0.9) +
  scale_y_continuous() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(), legend.title = element_blank(), axis.title.x = element_blank()) +
  labs(
    title = "Forestry Service Calls by NYC Borough",
    subtitle = "From February 2015 to September 2021",
    y = "Total calls",
    color = "Boroughs",
    caption = "Source: NYC Parks"
    ) +
  scale_fill_brewer(palette = "Set2")
calls_boro_plot
```

### Plot: Types of Calls in all NYC
Creating a bar graph of total types of calls by borough.
```{r}
call_types_nyc_plot <- ggplot(call_types_boro, aes(year, total_calls, group = category, fill = category)) +
  geom_col(size= 1, alpha = 0.9) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(), legend.title = element_blank()) +
  labs(
    title = "Types of Forestry Service Calls in all NYC",
    subtitle = "From February 2015 to September 2021",
    x = "Year",
    y = "Total calls",
    color = "Boroughs",
    caption = "Source: NYC Parks"
    ) +
  scale_fill_brewer(palette = "Set2")
call_types_nyc_plot
```

### Plot: Types of Calls by Borough
Creating a bar graph of total types of calls for each borough.
```{r}
call_types_boro_plot <- ggplot(call_types_boro, aes(year, total_calls, group = category, fill = category)) +
  geom_col(size = 1, alpha = 0.9) +
  scale_y_continuous() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(), legend.title = element_blank(), axis.title.x = element_blank()) +
  labs(
    title = "Types of Calls by Borough",
    subtitle = "From February 2015 to September 2021",
    x = "Year",
    y = "Total calls",
    color = "Boroughs",
    caption = "Source: NYC Parks"
  ) +
  scale_fill_brewer(palette = "Set2") +
  facet_wrap(~boro_name)
call_types_boro_plot
```

### Analysis: Calls by Community Board
Finding total number of calls and average of canopy cover by community board, excluding open cases, by filtering out non-community board values and grouping by district and borough.
```{r}
calls_cb <- calls %>% 
  filter(cb_name != "MN64") %>%  
  filter(cb_name != "BX26") %>%  
  filter(cb_name != "BX27") %>%  
  filter(cb_name != "BX28") %>%  
  filter(cb_name != "BK55") %>%  
  filter(cb_name != "BK56") %>% 
  filter(cb_name != "QN80") %>% 
  filter(cb_name != "QN81") %>% 
  filter(cb_name != "QN82") %>% 
  filter(cb_name != "QN83") %>% 
  filter(cb_name != "QN84") %>% 
  filter(cb_name != "SI95") %>% 
  group_by(cb_name, boro_name) %>% 
  summarise(
    total_calls = n(),
    canopy_cover = mean(canopy_cover, na.rm = TRUE),
    ) %>% 
  arrange(desc(total_calls)) %>% 
  ungroup()
calls_cb
```

### Analysis: Types of calls by Community Board
Finding total number of types of calls by community board, excluding open cases, by grouping by year, community board, borough, and category, and sorting by number of calls.
```{r}
call_types_cb <- calls %>% 
  mutate(year = str_sub(start, 1, 4)) %>% 
  group_by(year, cb_name, boro_name, category) %>% 
  summarise(total_calls = n()) %>% 
  arrange(desc(total_calls)) %>% 
  ungroup()
call_types_cb
```

### Plot: Ranking of Service Calls by Community Boards
Creating a bar graph of total calls by community boards.
```{r}
calls_cb_plot <- 
  ggplot(calls_cb, aes(reorder(cb_name, total_calls), total_calls, fill = boro_name)) +
  geom_col(size= 3, alpha = 0.9) +
  coord_flip() +
  scale_y_continuous() +
  theme_fivethirtyeight() +
  theme(
    axis.title = element_text(), 
    legend.title = element_blank(), 
    axis.title.y = element_blank()
    ) +
  labs(
    title = "Forestry Service Calls by NYC Community Board",
    subtitle = "From February 2015 to September 2021",
    x = "Community Boards",
    y = "Total calls",
    caption = "Source: NYC Parks"
  ) +
  scale_fill_brewer(palette = "Set2")
calls_cb_plot
```

### Plot: Ranking of Tree Canopy Cover by Community Board
Creating a bar graph of total tree canopy coverage by community boards.
```{r}
canopy_cb_plot <- 
  ggplot(calls_cb, aes(reorder(cb_name, canopy_cover), canopy_cover, fill = boro_name)) +
  geom_col(size= 3, alpha = 0.9) +
  coord_flip() +
  scale_y_continuous() +
  theme_fivethirtyeight() +
  theme(
    axis.title = element_text(), 
    legend.title = element_blank(), 
    axis.title.y = element_blank()
    ) +
  labs(
    title = "Tree Canopy Cover by NYC Community Board",
    subtitle = "From February 2015 to September 2021",
    x = "Community Boards",
    y = "Total proportion of tree canopy coverage",
    caption = "Source: New Yorkers for Parks"
  ) +
  scale_fill_brewer(palette = "Set2")
canopy_cb_plot
```

### Plot: Tree Canopy Coverage vs. Forestry Service Calls
Creating a scatterplot of tree canopy coverage and total calls by community board. 
```{r}
calls_vs_canopy <-
  ggplot(calls_cb, aes(total_calls, canopy_cover, color = boro_name, label = cb_name)) +
  geom_point(size= 3, alpha = 0.7, stroke = 0) +
  geom_text(aes(label=ifelse(total_calls>22000,as.character(cb_name),'')),hjust=-0.1,vjust=-0.1) +
  geom_text(aes(label=ifelse(canopy_cover>0.31,as.character(cb_name),'')),hjust=-0.1,vjust=-0.1) +
  theme_fivethirtyeight() +
  theme(
    axis.title = element_text(), 
    legend.title = element_blank()
    ) +
  labs(
    title = "Tree Canopy Coverage vs. Forestry Service Calls",
    subtitle = "By NYC Community Board",
    x = "Total calls since 2015",
    y = "Proportion of canopy cover",
    caption = "Source: NYC Parks and New Yorkers for Parks"
  ) +
  scale_color_brewer(palette = "Set2")
calls_vs_canopy
```

